Storing messages with multipartite neural cliques

### 2014-07-21 Articles/article.tex
...

        Note that there are $\binom{\ell}{a}$ possible choices for $a$ activations among $\ell$ units, giving the following correspondance:
\[
\binom{\ell}{a} = | \mathcal{A} |\;.
\]
Without loss of generality, we therefore consider that $\mathcal{A}$ is composed of the list of all binary vectors containing exactly $a$ ones that we index from 1 to $\binom{\ell}{a}$: $\mathcal{A} = \{\mathcal{A}_1,\dots,\mathcal{A}_{\binom{\ell}{a}}\}$.

@@

	Formally, we can describe the network using the adjacency matrix $W$ of the underlying graph. By construction this matrix is symmetric and binary. For a message $\mu$, note that $\mu \mu^{\intercal}$ is the cooccurence matrix associated with $\mu$. Before defining $W$, we first introduce the following matrix:
\[
	\overline{W} = \max_{\mu\in \mathcal{M}} \mu \mu^{\intercal} %% = \bigvee_{x\in \mathcal{M}} x x^{\intercal}
        \;,
\]
where $\max$ is a coefficient-to-coefficient max function. $W$ is obtained from $\overline{W}$ by removing all edges connecting units in a same cluster. We obtain a binary symmetric matrix $W$ with null blocks on the diagonal. The reason why we remove intracluster connections will be discussed in the following section.

@@

	We now study the performance of the new retrieving rule on distinct problems of retrieval : first partially erased messages, then erroneous messages.%, finally partially erased messages when the network has been damaged.

	Messages will be considered as independent and identically distributed according to the uniform distribution. We compute theoretical error rates after one iteration of the retrieving rule. We compare those results to empirical error rates obtained through simulations\footnote{Simulations are done by taking means on pool of five networks and messages sets, and each point is drawn for a different network and messages set.}. We also compare efficiencies of distinct networks for distinct values of the parameter $a$ (weight of the constant weight code) in the sense of information theory and show improvements over the case where $a = 1$.

	We first compute the theoretical density of the network, that is to mean the probability for a random edge $(i, i')$ to be present in the network (i.e. $W_{ii'} = 1$). We then express the theoretical error rate after one iteration of the retrieving rule in function of the density. All this computations (on error rates) are done assuming existences of edges between random units are independent, which is false but is a convenient approximation that gives results matching very closely results from simulations.

@@

@@

		Theoretical and empirical densities match very closely, see Figure \ref{densiteth}. %gap of theoretical density with density in simulations : under $1$ percent,

@@

	\begin{figure}[!htb]
		\includegraphics[width=8.5cm]{Courbes/remplacement_densite_c8l256a4}
		\caption{Theoretical and empirical densities for a network of 8 clusters, 256 units per cluster and 4 activated units per cluster}
			\label{densiteth}
	\end{figure}

	\begin{figure}[!htb]
		\includegraphics[width=8.5cm]{Courbes/remplacement_figure2g1} %{comparaison_regles_pool5_a2c4l512e2}
		\caption{Evolution of error rate for increasing number of stored messages, 2 activities per cluster, 4 clusters, 512 units per cluster, two erasures, each point is the mean of 5 networks with 1000 sampled messages, analytical result (for one iteration) in continuous black}
			\label{erasuresth}
		\end{figure}

@@

		$\eta_m = P_{retrieve}  \frac{2 M\left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2} \times \left [\sum_{k = c_e+1}^{c} \left (1-\frac{1}{{\ell \choose a}} \right )^k \left (\frac{1}{{ell \choose a}} \right )^{c-k} \right ] ^{-(M-1)}$

@@

	With random noise on edges, hardness of the problem must take into account the capacity of the binary symmetric channel (the closer $\psi$ is of $0.5$, the harder the problem is). Our new is efficiency is written as :  	
	$\eta_{m, \psi} = P_{retrieve} m  \frac{2 \left(c \log_2{\ell \choose a } - \log_2(m) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times (1-{\ell \choose a}^{c - c_e})^{-(m-1)} \times \frac{1}{1 + \psi \log_2(\psi) +(1-\psi) \log_2(1-\psi)} $

@@

	Figure \ref{comppsi} seems to indicate a linear relationship (after a transition for multiple activities) between $\eta_\psi -\eta_0$ and $\psi$.
	\section{Maximum theoretical efficiency}
		It was shown in \cite{Palm1980} that maximal asymptotic capacity of Willshaw networks is $\log 2$. We adapt this result to multipartite clique clustered networks.

>>>>

        Note that there are $\binom{\ell}{a}$ possible choices for $a$ activations among $\ell$ units, giving the following correspondence:
\[
\binom{\ell}{a} = | \mathcal{A} |\;.
\]
Without loss of generality, we therefore consider that $\mathcal{A}$ is composed of the list of all binary vectors containing exactly $a$ ones that we index from 1 to $\binom{\ell}{a}$: $\mathcal{A} = \{\mathcal{A}_1,\dots,\mathcal{A}_{\binom{\ell}{a}}\}$.

@@

	Formally, we can describe the network using the adjacency matrix $W$ of the underlying graph. By construction this matrix is symmetric and binary. For a message $\mu$, note that $\mu \mu^{\intercal}$ is the co-occurrence matrix associated with $\mu$. Before defining $W$, we first introduce the following matrix:
\[
	\overline{W} = \max_{\mu\in \mathcal{M}} \mu \mu^{\intercal} %% = \bigvee_{x\in \mathcal{M}} x x^{\intercal}
        \;,
\]
where $\max$ is a coefficient-to-coefficient max function. $W$ is obtained from $\overline{W}$ by removing all edges connecting units in a same cluster. We obtain a binary symmetric matrix $W$ with null blocks on the diagonal. The reason why we remove intracluster connections will be discussed in the following section.

@@

	We now describe some static properties of the networks as the density of edges.

	\section{Static properties of multipartite clique networks}

	We study the properties of the network after storing messages in it. We are interested in how much edges have been added to the network and if the process introduced confusion between messages.

	Messages stored in the network will be considered independent and identically distributed according to the uniform distribution. This randomization allows us to compute the theoretical density of the network. %under the approximation that edges are independent. 
	Density is a crucial parameter of the network that can be used to predict retrieval rates for different problems. Moreover we compute asymptotic efficiency thanks to the formula for density.

	\subsection{Density}

	The density $d$ of the network is the proportion of edges present in the network : $d = \frac{\# \{(i, i') \} \, | W_{ii'} = 1 \}}{c(c-1)\ell^2}$.

	\subsubsection{Theoretical density}

	We first compute the theoretical density of the network, that is to mean the probability for a random edge $(i, i')$ to be present in the network (i.e. $W_{ii'} = 1$).

@@

	Theoretical and empirical densities match very closely, see Figure \ref{densiteth}. %gap of theoretical density with density in simulations : under $1$ percent, 
	\begin{figure}[!htb]
		\includegraphics[width=8.5cm]{Courbes/remplacement_densite_c8l256a4}
		\caption{Theoretical and empirical densities for a network of 8 clusters, 256 units per cluster and 4 activated units per cluster}
			\label{densiteth}
	\end{figure}

	\subsection{Accepting random messages}

	\section{Maximum theoretical efficiency}
		It was shown in \cite{Palm1980} that maximal asymptotic capacity of Willshaw networks is $\log 2$. We adapt this result to multipartite clique clustered networks.

	We now describe the distinct problems of messages retrieval that we study the performance of.

	\section{Problems statements}

	\subsection{Erasures}

	\subsection{Random replacements}

	\subsection{Resilience}

	We now study the performance of the new retrieving rule on those distinct problems of retrieval : first partially erased messages, then erroneous messages.%, finally partially erased messages when the network has been damaged.

	Thanks to the theoretical formula for density, we compute analytical error rates after one iteration of the retrieving rule. We compare those results to empirical error rates obtained through simulations\footnote{Simulations are done by taking means on pool of five networks and messages sets, and each point is drawn for a different network and messages set.}. We also compare efficiencies of distinct networks for distinct values of the parameter $a$ (weight of the constant weight code) in the sense of information theory and show improvements over the case where $a = 1$.

	We then express the theoretical error rate after one iteration of the retrieving rule in function of the density. All this computations (on error rates) are done assuming existences of edges between random units are independent, which is false but is a convenient approximation that gives results matching very closely results from simulations.

@@

@@

%	\begin{figure}[!htb]
%		\includegraphics[width=8.5cm]{Courbes/remplacement_figure2g1} %{comparaison_regles_pool5_a2c4l512e2}
%		\caption{Evolution of error rate for increasing number of stored messages, 2 activities per cluster, 4 clusters, 512 units per cluster, two erasures, each point is the mean of 5 networks with 1000 sampled messages, analytical result (for one iteration) in continuous black}
%			\label{erasuresth}
%		\end{figure}

		\begin{figure}[!htb]
		\includegraphics[width=8.5cm]{Courbes/fig3c4l512e2a2} %{comparaison_regles_pool5_a2c4l512e2}
		\caption{Evolution of error rate for increasing number of stored messages, 2 activities per cluster, 4 clusters, 512 units per cluster, two erasures, each point is the mean of 5 networks with 1000 sampled messages, analytical result (for one iteration) in continuous black}
			\label{erasuresth}
		\end{figure}

@@

		$\eta_m = P_{retrieve}  \frac{2 M\left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2} \times \left [\sum_{k = c_e+1}^{c} \left (1-\frac{1}{{\ell \choose a}} \right )^k \left (\frac{1}{{\ell \choose a}} \right )^{c-k} \right ] ^{-(M-1)}$

@@

	With random noise on edges, hardness of the problem must take into account the capacity of the binary symmetric channel (the closer $\psi$ is of $0.5$, the harder the problem is). Our new is efficiency is written as :  	
	$\eta_{m, \psi} = P_{retrieve}  \frac{2 M \left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)l^2} \times (1-{\ell \choose a}^{c - c_e})^{-(M-1)} \times \frac{1}{1 + \psi \log_2(\psi) +(1-\psi) \log_2(1-\psi)} $

@@

	Figure \ref{comppsi} seems to indicate a linear relationship (after a transition for multiple activities) between $\eta_\psi -\eta_0$ and $\psi$.
